Highly Available Transactions: Virtues and Limitations

### 2013-04-23 writeup/conclusion.tex
editing pass

\section{Conclusion}
The current state of database software offers uncomfortable and
unnecessary choices between availability and transactional semantics.
We show, however, through analysis and experiments, that the
availability goals of the NoSQL movement will remain a critical aspect
of many future data storage systems. Accordingly, we expose a broad
design space of Highly Available Transactional Systems (HATs), which
can offer the key benefits of highly available distributed
systems---''always on'' operation during partitions and low-latency
operations---while also providing a family of transactional isolation
levels and replica semantics that have been adopted in practice.  In
taxonomizing this design space, we show that many transactional
guarantees are achievable with high availability, while a few, like
preventing Lost Update and Write Skew, are not. From a research
perspective, this offers a long-overdue connection between storage
semantics studied in the database and distributed systems communities.

With the above understanding in hand, the next challenge is to return
to issues of performance and systems architecture within the HAT
design space.  Based on our understanding of what is desired in the
field and newfound knowledge of what is possible to achieve, we
believe this is an area of great opportunity in the coming years.

>>>>

\section{Conclusion}
The current state of database software offers uncomfortable and
unnecessary choices between availability and transactional semantics.
However, through our analysis and experiments, we have demonstrated
that the NoSQL movement's goals of high availability will remain a
critical aspect of many future data storage systems. Accordingly, we
expose a broad design space of Highly Available Transactions (HATs),
which can offer the key benefits of highly available distributed
systems---''always on'' operation during partitions and low-latency
operations---while also providing a family of transactional isolation
levels and replica semantics that have been adopted in practice.  In
taxonomizing this design space, we show that many transactional
guarantees are achievable with high availability, while a few, like
preventing Lost Update and Write Skew, are not. From a research
perspective, this offers a long-overdue connection between storage
semantics studied in the database and distributed systems communities.

With the above understanding in hand, the next challenge is to return
to issues of performance and systems architecture within the HAT
design space.  Based on our understanding of what is desired in the
field and newfound knowledge of what is possible to achieve, we
believe HATs will be an area of great opportunity in the coming years.

### 2013-04-23 writeup/evaluation.tex
editing pass

\begin{myenumerate}
\item We revisit traditional database concurrency control with a focus
  on coordination costs and on high availability.
\item We examine the properties required by a realistic OLTP
  application based on the TPC-C benchmark.
\item We perform a brief experimental evaluation of HAT versus non-HAT
  properties on public cloud infrastructure.
\end{myenumerate}

@@

\vspace{.5em}\noindent\textbf{Serializability} To establish a serial
order on transactions, algorithms for achieving serializability of
general-purpose read-write transactions in a distributed
setting~\cite{bernstein-book} may require at least one round trip time
(RTT) before committing. As an example, traditional two-phase locking
for a transaction of length $T$ may require $T$ \texttt{lock}
operations and will require at least one \texttt{lock} and
\texttt{unlock} operation.  In a distributed environment, each of
these operations requires coordination, either with other database
replicas or with a lock service. If this coordination mechanism is
unavailable, transactions cannot safely commit. Similarly, optimistic
concurrency control requires coordinating via a validation step, while
deterministic transaction scheduling~\cite{deterministic-scheduling}
requires a scheduler. Serializability under multi-version concurrency
control also requires checking for update conflicts. The reliance on a
globally agreed total order necessitates a minimum of one round-trip
to a designated master or coordination service for each of these
classic algorithms.  The cost of the (minimum one) round trip will be
determined by the deployment environment, as we saw in
Section~\ref{sec:motivation}; we will demonstrate this cost on public
cloud infrastructure in the next section.

\vspace{.5em}\noindent\textbf{Non-serializability} Many existing
distributed implementations of weak isolation are not highly
available. Lock-based proposals such as those used to provide weak
isolation in Gray's original proposal~\cite{gray-isolation} do not
degrade gracefully in the presence of partial failures. (Note, however,
that lock-based protocols \textit{do} offer the benefit of recency
guarantees.) While multi-versioned storage systems allow for a variety
of transactional guarantees, we have not seen proposals for truly weak
isolation (e.g., non-``tentative update'' schemes) in this context.
The MDCC~\cite{mdcc} protocol offers Read Committed isolation with
Lost Update avoidance but is similarly unavailable due to its reliance
on preventing write conflicts. Chan and Gray's read-only transactions
provide read-only transactions with item-cut isolation, causal
consistency, and transactional atomicity (session PL-2L~\cite{adya})
but are unavailable in the presence of coordinator
failure~\cite{readonly}, similar to read-only and write-only
transactions more recently proposed by Eiger~\cite{eiger}; Causal
Serializability offers a similar model (with unavailable
implementation): causal consistency with a variant of Read Uncommitted
between transactions that write to the same data
item~\cite{raynal-causal}.  Bolt-on causal consistency~\cite{bolton},
Brantner's S3 database~\cite{kraska-s3}, and
Bayou~\cite{sessionguarantees} can all provide variants of session
PL-2L with high availability, but none provide this HAT functionality
without substantial modification; Swift~\cite{swift} is closest to
providing full HAT operation in a sticky context. As  seen, it
is possible to implement many guarantees weaker than
serializability---including guarantees that are achievable with high
availability---and still not achieve high availability.

@@

\vspace{.5em}\noindent\textbf{Commutativity and Monotonicity} Recent
work on the CALM Theorem~\cite{calm} and Commutative and Replicated Data
Types~\cite{crdt} demonstrates that, if updates logically commute,
then they can often be safely performed in different orders at
different replicas. Accordingly, as long as all writes are delivered
to all replicas, then a system executing monotonic logic with
commutative operators may not suffer from application-level
consistency anomalies as a result of Lost Update or Write Skew
anomalies. However, applications with non-monotonic state mutation
will not, in general, be able to maintain application-level
consistency constraints with HATs alone. Moreover, applications with
requiring bounded update visibility latency should opt for
unavailability.

\vspace{.5em}\noindent\textbf{TPC-C} To better understand the impact
of HAT-compliance in an application context, we consider a concrete
application: the TPC-C benchmark. Surprisingly, four of five
transactions can be executed with HATs, while the fifth may require
unavailability.

@@

New-Order presents two challenges: ID assignment and stock
maintenance. First, each New-Order transaction requires a unique ID
number for the order. We can create a unique number by, say,
concatenating the client ID and a timestamp. However, the TPC-C
specification requires order numbers to be \textit{sequentially}
assigned within a district, which requires preventing Lost
Update. Accordingly, HATs cannot provide compliant TPC-C execution but
can maintain uniqueness constraints. Second, the New-Order transaction
decrements inventory counts: what if the count becomes negative?
Fortunately, TPC-C New-Order restocks each item's inventory count
(increments by 91) if it would become negative as the result of
placing an order. This means that, even in the presence of concurrent
New-Orders, an item's stock will never fall below zero. This is TPC-C
compliant, but a HAT system might end up with much more stock
than in a non-HAT-compliant implementation.

Delivery is challenging due to non-monotonicity. Each Delivery deletes
a pending order from the New-Order Table and should be idempotent in
order to avoid billing a customer twice; this implies a need to
prevent Lost Update. This issue  can be avoided  by moving the
non-monotonicity to the real world---the carrier that picks up the
package for an order can ensure that she is the only carrier who has
done so---but cannot provide a correct execution with HATs
alone. However, according to distributed transaction
architects~\cite{entitygroup}, these compensatory actions are
relatively common in real-world business processes.

Throughout execution, TPC-C also requires the maintenance of several
integrity constraints. For example, Consistency Condition 1 (3.3.2.1)
requires that each warehouse's sales count must reflect the sum of its
subordinate sales districts. This integrity constraint spans two
tables but, given the ability to update rows in both tables atomically
via transactional atomicity, can be easily maintained. Consistency
Conditions 4 through 12 (3.3.2.4-12) can similarly be satisfied by
applying updates atomically across tables. Consistency Conditions 2
and 3 (3.3.2.2-3) concern order ID assignment and are more
problematic, as we have discussed.  Finally, while TPC-C is not
subject to multi-key anomalies, we note that many TPC-E isolation
tests, such as one that requires simultaneously modifying a product
description and its image, are also passable using HATs.

>>>>

\begin{myenumerate}\vspace{-.5em}
\item We revisit traditional database concurrency control with a focus
  on coordination costs and on high availability.
\item We examine the properties required by a realistic OLTP
  application based on the TPC-C benchmark.
\item We perform a brief experimental evaluation of HAT versus non-HAT
  properties on public cloud infrastructure.
\end{myenumerate}\vspace{-.5em}

@@

\vspace{.5em}\noindent\textbf{Serializability} To establish a serial
order on transactions, algorithms for achieving serializability of
general-purpose read-write transactions in a distributed
setting~\cite{bernstein-book} can require at least one round trip time
(RTT) before committing. As an example, traditional two-phase locking
for a transaction of length $T$ can require $T$ \texttt{lock}
operations and will require at least one \texttt{lock} and
\texttt{unlock} operation.  In a distributed environment, each of
these operations requires coordination, either with other database
servers or with a lock service. If this coordination mechanism is
unavailable, transactions cannot safely commit. Similarly, optimistic
concurrency control requires coordinating via a validation step, while
deterministic transaction scheduling~\cite{deterministic-scheduling}
requires a scheduler. Serializability under multi-version concurrency
control also requires checking for update conflicts. The reliance on a
globally agreed total order necessitates a minimum of one round-trip
to a designated master or coordination service for each of these
classic algorithms.  The cost of round trips will be determined by the
deployment environment, as we saw in Section~\ref{sec:motivation}; we
will demonstrate this cost on public cloud infrastructure in the next
section.

\vspace{.5em}\noindent\textbf{Non-serializability} Many existing
distributed implementations of weak isolation are not highly
available. Lock-based proposals such as those used to provide weak
isolation in Gray's original proposal~\cite{gray-isolation} do not
degrade gracefully in the presence of partial failures. (Note,
however, that lock-based protocols \textit{do} offer the benefit of
recency guarantees.) While multi-versioned storage systems allow for a
variety of transactional guarantees, we have not seen proposals for
truly weak isolation (e.g., non-``tentative update'' schemes) in this
context.  The MDCC~\cite{mdcc} protocol offers Read Committed
isolation with Lost Update avoidance but is similarly unavailable due
to its reliance on preventing write conflicts. Chan and Gray's
read-only transactions provide read-only transactions with item-cut
isolation, causal consistency, and transactional atomicity (session
PL-2L~\cite{adya}) but are unavailable in the presence of coordinator
failure~\cite{readonly}, similar to read-only and write-only
transactions more recently proposed by Eiger~\cite{eiger}; Causal
Serializability offers a similar model (with unavailable
implementation): causal consistency with a variant of Read Uncommitted
between transactions that write to the same data
item~\cite{raynal-causal}.  Brantner's S3 database~\cite{kraska-s3}
and Bayou~\cite{sessionguarantees} can all provide variants of session
PL-2L with high availability, but none provide this HAT functionality
without substantial modification; Swift~\cite{swift} and bolt-on
causal consistency~\cite{bolton} are closest to providing full HAT
semantics operation in a sticky context. As seen, it is possible to
implement many guarantees weaker than serializability---including
guarantees that are achievable with high availability---and still not
achieve high availability.

@@

\vspace{.5em}\noindent\textbf{Commutativity and Monotonicity} Recent
work on the CALM Theorem~\cite{calm} and Commutative and Replicated
Data Types~\cite{crdt} demonstrates that, if updates logically
commute, then they can often be safely performed in different orders
at different replicas. Accordingly, as long as all writes are
delivered to all replicas, then a system executing monotonic logic
with commutative operators may not suffer from application-level
consistency anomalies as a result of Lost Update or Write Skew
anomalies. However, applications with non-monotonic state mutation
will not, in general, be able to maintain application-level
consistency constraints with HATs alone--in particular, applications
requiring bounded update visibility latency should opt for
unavailability.

\vspace{.5em}\noindent\textbf{TPC-C} To better understand the impact
of HAT-compliance in an application context, we consider a concrete
application: the TPC-C benchmark~\cite{tpcc}. Surprisingly, four of
five transactions can be executed with HATs, while the fifth may
require unavailability.

@@

New-Order presents two challenges: ID assignment and stock
maintenance. First, each New-Order transaction requires a unique ID
number for the order. We can create a unique number by, say,
concatenating the client ID and a timestamp. However, the TPC-C
specification requires order numbers to be \textit{sequentially}
assigned within a district, which requires preventing Lost
Update. Accordingly, HATs cannot provide compliant TPC-C execution but
can stil maintain uniqueness constraints. Second, the New-Order transaction
decrements inventory counts: what if the count becomes negative?
Fortunately, TPC-C New-Order restocks each item's inventory count
(increments by 91) if it would become negative as the result of
placing an order. This means that, even in the presence of concurrent
New-Orders, an item's stock will never fall below zero. This is TPC-C
compliant, but a HAT system might end up with more stock
than in a non-HAT-compliant implementation.

Delivery is challenging due to non-monotonicity. Each Delivery deletes
a pending order from the New-Order table and should be idempotent in
order to avoid billing a customer twice; this implies a need to
prevent Lost Update. This issue  can be avoided  by moving the
non-monotonicity to the real world---the carrier that picks up the
package for an order can ensure that she is the only carrier who has
done so---but cannot provide a correct execution with HATs
alone. However, according to distributed transaction
architects~\cite{entitygroup}, these compensatory actions are
relatively common in real-world business processes.

Throughout execution, TPC-C also requires the maintenance of several
integrity constraints. For example, Consistency Condition 1 (3.3.2.1)
requires that each warehouse's sales count must reflect the sum of its
subordinate sales districts. This integrity constraint spans two
tables but, given the ability to update rows in both tables atomically
via TA, can be easily maintained. Consistency Conditions 4 through 12
(3.3.2.4-12) can similarly be satisfied by applying updates atomically
across tables. Consistency Conditions 2 and 3 (3.3.2.2-3) concern
order ID assignment and are problematic. Finally, while TPC-C is not
subject to multi-key anomalies, we note that many TPC-E isolation
tests (i.e., simultaneously modifying a product description and its
thumbnail) are also achievable using HATs.
